Topological Equivariance Artist Model

### 2021-03-15 paper/paper.tex
q

%% A teaser figure can be included as follows
\teaser{
  \centering
  \includegraphics[width=\linewidth]{teaser.png}
  \caption{this is our model basically...}
  \label{fig:teaser}
}

@@

The contribution of this work is
\begin{description}
  \item[Topological] topology preserving relationship between data and graphic via continuous maps
  \item[Equivariant] property preservation from data component to visual representation as equivariant maps that carry a homomorphism of monoid actions
  \item[Artist] functional oriented visualization tool architecture built on the mathematical model to demonstrate the utility of the model
  \item [Model] prototype of the architecture built on Matplotlib's infrastructure to demonstrate the feasibility of the model
\end{description}

@@

\subsubsection{Fiber Space: Variables}
\label{sec:math:data:fiber}
\begin{figure}[tb]
  \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
  \includegraphics[width=\columnwidth]{fiber.png}
  \caption{These two datasets have the same base space \dbase\ of discrete points, but the square is a visual representation of the fiber space  $\dfiber=\reals\times\reals$, while the cube is a representation of $\dfiber=\reals\times\realsp\times\reals$}
  \label{fig:data_fiber}
\end{figure}

To formalize the structure of the data components, we use notation introduced by Spivak \cite{spivakSIMPLICIALDATABASES} that binds the components of the fiber to variable names. Spivak constructs a set \ftotal\ that is the disjoint union of all possible objects of types $\{\ftype_0, \ldots, \ftype_m\} \in \ftypes$, where \ftypes\ are the data types of the variables in the dataset. He then defines the single variable set \fttype\ 
\begin{equation}
    \label{eq:data_types}
    \begin{tikzcd}
        \fttype \arrow[r] \arrow[d, "\pi_{\fsection}"'] & \ftotal \arrow[d, "\pi"] \\
        \fnames \arrow[r, "\fsection"']                          & \ftypes       
    \end{tikzcd}
\end{equation}
which is \ftotal\ restricted to objects of type \ftype\ bound to variable name \fname. The \fttype\ lookup is by name to specify that every component is distinct, since multiple components can have the same type \ftype. Given \fsection, the fiber for a one variable dataset is
\begin{equation}
    \dfiber = \ftotal_{\fsection(\fname)} = \ftotal_{\ftype} 
\end{equation}
where \fsection\ is the schema binding variable name \fname\ to its datatype \ftype. A dataset with multiple variables has a fiber that is the cartesian cross product of $\ftotal_{\fsection}$ applied to all the columns:
\begin{equation}
F = \ftotal_{\fsection(\fname_{1})}\times \ldots \ftotal_{\fsection(\fname_{i})} \ldots\times \ftotal_{\fsection(\fname_{n})}
\end{equation}
which is equivalent to 
\begin{equation}
    \label{eq:math:data:fiber:decompose}
    \dfiber= \dfiber_{0} \times \ldots \times \dfiber_{i}\times\ldots\times \dfiber_{n}
\end{equation}
which allows us to decouple \dfiber\ into components $\dfiber_i$. Each component of \dfiber\ is a dimension of the topological fiber space, as illustrated in \autoref{fig:data_fiber}. The space $\dfiber=\reals\times\reals$ encodes that the dataset has two quantative continuous variables, while $\dfiber=\reals\times\reals\realsp$ encodes a dataset with three quantative contunous variables, one of which must be positive.

@@

 \begin{figure}[tb]
  \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
  \includegraphics[width=\columnwidth]{base.png}
  \caption{These two datasets have the same fiber space  $\dfiber=\reals\times\reals$, but the left dataset has a discrete continuity while the right is 1D continuous over the interval $\left[0,1\right]$.}
  \label{fig:data_continuity}
 \end{figure}

 As illustrated in ~\autoref{fig:data_continuity}, \dbase\ can have any number of dimensions, can be continuous or discrete, and is somewhat independent of the dimensions of the fiber. As with fibers and monoids, we can decompose the total space into components $\pi:\dtotal_i\rightarrow \dbase$ where
 \begin{equation}
    \label{eq:math:data:base:decompose}
     \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
 \end{equation}
 which is a decomposition of \dfiber. The \dbase\ remains the same because the connectivity of records does not change just because there are fewer elements in each record. By encoding this continuity in the model as \dbase\, the data model now explicitly carries information about its structure such that the implicit assumptions of the visualization algorithms are now explicit. The explicit topology is a concise way of distinguishing visualizations that appear identical, for example heatmaps and images.

 \subsubsection{Section: Values}
 \label{sec:math:data:section}
 \begin{figure}[tb]
  \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
  \includegraphics[width=\columnwidth]{fiberbundle.png}
  \caption{Each section in the fiber bundle is a unique continuous map from base space to fiber encoding the set of records in the dataset. The two sections, $\dsection^{(1)}$ and $\dsection^{(2)}$, shown here are in the set of global sections $\Gamma(\dtotal)$.}
  \label{fig:data_section}
 \end{figure}

 While the projection function $\pi:\dtotal \rightarrow\dbase$ ties together the base space \dbase\ with the fiber \dfiber, a section $\dsection: \dbase\rightarrow \dtotal$ encodes a dataset. A section function takes as input location $\dbasepoint \in \dbase$ and returns a record $\delement \in \dtotal$. For any fiber bundle, there exists a map
 \begin{equation}
     \begin{tikzcd}
         \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
                           & \dbase \arrow[u, "\dsection"', bend right]
     \end{tikzcd}
 \end{equation}
  such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$. The set of all global sections is denoted as $\Gamma(\dtotal)$. As illustrated in \autoref{fig:data_section}, the section is a continuous mapping from a location $\dbasepoint \in \dbase$ on the base space to a record $\delement \in \dfiber$ in the fiber. Assuming a trivial fiber bundle $\dtotal = \dbase \times \dfiber$, the section is 
 \begin{equation}
     \label{eq:section_return}
     \dsection(\dbasepoint) = (\dbasepoint, (g_{\dfiber_{0}}(\dbasepoint), \ldots, g_{\dfiber_{n}}(\dbasepoint)))
 \end{equation}
 where $g: \dbase \rightarrow \dfiber$ is the index function into the fiber component. This formulation of the section also holds on locally trivial sections of a non-trivial fiber bundle. As with \autoref{eq:math:data:fiber:decompose} and \autoref{eq:math:data:base:decompose}, \dsection\ can be decomposed into components 
 \begin{equation}
 \label{eq:math:data:section:decompose}
 \dsection= (\dsection_0,\ldots, \dsection_i, \dots, \dsection_n) 
 \end{equation}
 where each section $\dsection_i$ maps into a record on a component $\dfiber_i \in \dfiber$. This allows for accessing the data component wise in addition to accessing the data in terms of its location over \dbase.

@@

\subsubsection{Equivariant Topology: Graphic Base Space}
\label{sec:math:graphic:base}
\begin{figure}[tb]
  \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
  \includegraphics[width=\columnwidth]{retraction_maps.png}
  \caption{The 0D scatter \dbasepoint\ and 1D line \dbasepoint\ are thickened into \gbase\ with coordinates $\gbasepoint=(\gx, \gy)$ that are a region in an idealized 2D screen. The image has the same dimension in \gbase\ as in \dbase.}
  \label{fig:graphic_base}
 \end{figure}

Just as the \dbase\ encodes the connectivity of the records in the data, we propose an equivalent \gbase\ that encodes the connectivity of the rendered elements of the graphic. Formally, we require that \dbase\ be a deformation retract\cite{RetractionTopology2020} of \gbase\ so that \dbase\ and \gbase\ have the same homotopy. The surjective map $\vindex: \gbase \rightarrow \dbase$ 
\begin{equation}
    \begin{tikzcd}
        \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
        \dbase                   & \gbase \arrow[l, "\vindex"']
    \end{tikzcd}
\end{equation}
goes from region $\gbasepoint \in \gbase_{\dbasepoint}$ to its associated point $\gbasepoint$. While \gbase\ must have the same continuity as \dbase\, it is sometimes the  thickened version shown in \autoref{fig:graphic_base}. This thickening is necessary when the dimensionality of \dbase\ is less than the dimensionality of the target display. For example, a \dbasepoint\ that is 0D in \dbase\ cannot be represented on screen unless it is thickened to 2D to encode the connectivity of the points in \gfiber\ that visually represent the record at \dbasepoint.

@@

\subsubsection{Equivariant Visual Component Maps}
\label{sec:math:artist:nu}
\begin{figure}[tb]
  \centering
  \includegraphics[width=\columnwidth]{equivariance_nu.png}
  \caption{In this artist, \vchannel\ maps the strings to the emojis. This \vchannel\ is equivariant because the monoid actions, represented by the colored arrows, are the same on the input \dsection\ and output \vsection\ sets.}
  \label{fig:math:artist:nu}
\end{figure}
 We define the visual transformers \vchannel\ 
\begin{equation}
  \label{eq:math:artist:nu}
  \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
as the set of equivariant maps $\vchannel_i: \dsection_i \mapsto \vsection_i$. Given $\monoid_i$ is the monoid action on $\dtotal_i$ and that there is a monoid ${\monoid_i}^{\prime}$ on $\vtotal_i$, then there is a monoid homomorphism from $\varphi:\monoid_i \rightarrow {\monoid_i}^{\prime}$ that \vchannel\ must preserve. As mentioned in \autoref{sec:math:data:monoid}, monoid actions define the structure on the fiber components and are therefore the basis for equivariance. 
A validly constructed \vchannel\ is one where the diagram of the monoid transform $m$ commutes such that 
\begin{equation}
  \label{eq:math:artist:nu_commute}
\begin{tikzcd}
  \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
  \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
\end{tikzcd}
\end{equation}
In general, the data fiber $\dfiber_{i}$ cannot be assumed to be of the same type as the visual fiber $\vfiber_{i}$ and the actions of \monoid\ on $\dfiber_{i}$ cannot be assumed to be the same as the actions of $\monoid^{\prime}$ on \vfiber; therefore an equivariant $\vchannel_i$ must satisfy the constraint  
\begin{equation}
\vchannel_i(m_{\delement}(\dtotal_i)) = \varphi(m_{\delement})(\vchannel_i(\dtotal_i))
\end{equation} 
such that $\varphi$ maps a monoid action on data to a monoid action on visual elements. However, we can construct a monoid action of \monoid\ on $\vfiber_{i}$ that is compatible with a monoid action of \monoid\ on $\dfiber_{i}$. We can compose the monoid actions on the visual fiber $\monoid^{\prime} \times \vfiber_{i} \rightarrow \vfiber_{i}$ with the homomorphism $\varphi$ that takes \monoid\ to $\monoid^{\prime}$. This allows us to define a monoid action on \vfiber\ of \monoid\ that is $(m, \velement) \rightarrow \varphi(m)\bullet\velement$. Therefore, without a loss of generality, we can assume that an action of \monoid\ acts on $\dfiber_{i}$ and on $\vfiber_{i}$ compatibly such that $\varphi$ is the identity function.

@@

>>>>

%% A teaser figure can be included as follows
\teaser{
  \centering
  \includegraphics[width=\linewidth]{teaser.png}
  \caption{Visualizations consist of topologically equivariant maps. There is a set of monoid action equivariant maps from data components to visual components \vchannel\ that are then reduced via \vmark\ into a single graphic, and there is a deformation retraction \vindex\ from graphic continuity to data continuity.}
  \label{fig:teaser}
}

@@

The contribution of this work is
\begin{enumerate}
  \item topology preserving relationship between data and graphic via continuous maps
  \item property preservation from data component to visual representation as equivariant maps that carry a homomorphism of monoid actions
  \item functional oriented visualization tool architecture built on the mathematical model to demonstrate the utility of the model
  \item prototype of the architecture built on Matplotlib's infrastructure to demonstrate the feasibility of the model
\end{enumerate}

@@

\subsubsection{Fiber Space: Variables}
\label{sec:math:data:fiber}
\begin{figure}[htb]
  \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
  \includegraphics[width=\columnwidth]{fiber.png}
  \caption{These two datasets have the same base space \dbase\ of discrete points, but the square is a visual representation of the fiber space  $\dfiber=\reals\times\reals$, while the cube is a representation of $\dfiber=\reals\times\realsp\times\reals$}
  \label{fig:math:data:fiber}
\end{figure}

To formalize the structure of the data components, we use notation introduced by Spivak \cite{spivakSIMPLICIALDATABASES} that binds the components of the fiber to variable names. Spivak constructs a set \ftotal\ that is the disjoint union of all possible objects of types $\{\ftype_0, \ldots, \ftype_m\} \in \ftypes$, where \ftypes\ are the data types of the variables in the dataset. He then defines the single variable set \fttype\ 
\begin{equation}
    \label{eq:data_types}
    \begin{tikzcd}
        \fttype \arrow[r] \arrow[d, "\pi_{\fsection}"'] & \ftotal \arrow[d, "\pi"] \\
        \fnames \arrow[r, "\fsection"']                          & \ftypes       
    \end{tikzcd}
\end{equation}
which is \ftotal\ restricted to objects of type \ftype\ bound to variable name \fname. The \fttype\ lookup is by name to specify that every component is distinct, since multiple components can have the same type \ftype. Given \fsection, the fiber for a one variable dataset is
\begin{equation}
    \dfiber = \ftotal_{\fsection(\fname)} = \ftotal_{\ftype} 
\end{equation}
where \fsection\ is the schema binding variable name \fname\ to its datatype \ftype. A dataset with multiple variables has a fiber that is the cartesian cross product of $\ftotal_{\fsection}$ applied to all the columns:
\begin{equation}
F = \ftotal_{\fsection(\fname_{1})}\times \ldots \ftotal_{\fsection(\fname_{i})} \ldots\times \ftotal_{\fsection(\fname_{n})}
\end{equation}
which is equivalent to 
\begin{equation}
    \label{eq:math:data:fiber:decompose}
    \dfiber= \dfiber_{0} \times \ldots \times \dfiber_{i}\times\ldots\times \dfiber_{n}
\end{equation}
which allows us to decouple \dfiber\ into components $\dfiber_i$. Each component of \dfiber\ is a dimension of the topological fiber space, as illustrated in \autoref{fig:math:data:fiber}. The space $\dfiber=\reals\times\reals$ encodes that the dataset has two quantative continuous variables, while $\dfiber=\reals\times\reals\realsp$ encodes a dataset with three quantative contunous variables, one of which must be positive.

@@

 \begin{figure}[htb]
  \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
  \includegraphics[width=\columnwidth]{base.png}
  \caption{These two datasets have the same fiber space  $\dfiber=\reals\times\reals$, but the left dataset has a discrete continuity while the right is 1D continuous over the interval $\left[0,1\right]$.}
  \label{fig:math:data:base}
 \end{figure}

 As illustrated in ~\autoref{fig:math:data:base}, \dbase\ can have any number of dimensions, can be continuous or discrete, and is somewhat independent of the dimensions of the fiber. As with fibers and monoids, we can decompose the total space into components $\pi:\dtotal_i\rightarrow \dbase$ where
 \begin{equation}
    \label{eq:math:data:base:decompose}
     \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
 \end{equation}
 which is a decomposition of \dfiber. The \dbase\ remains the same because the connectivity of records does not change just because there are fewer elements in each record. By encoding this continuity in the model as \dbase\, the data model now explicitly carries information about its structure such that the implicit assumptions of the visualization algorithms are now explicit. The explicit topology is a concise way of distinguishing visualizations that appear identical, for example heatmaps and images.

 \subsubsection{Section: Values}
 \label{sec:math:data:section}
 \begin{figure}[htb]
  \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
  \includegraphics[width=\columnwidth]{fiberbundle.png}
  \caption{Each section in the fiber bundle is a unique continuous map from base space to fiber encoding the set of records in the dataset. The two sections, $\dsection^{(1)}$ and $\dsection^{(2)}$, shown here are in the set of global sections $\Gamma(\dtotal)$.}
  \label{fig:math:data:section}
 \end{figure}

 While the projection function $\pi:\dtotal \rightarrow\dbase$ ties together the base space \dbase\ with the fiber \dfiber, a section $\dsection: \dbase\rightarrow \dtotal$ encodes a dataset. A section function takes as input location $\dbasepoint \in \dbase$ and returns a record $\delement \in \dtotal$. For any fiber bundle, there exists a map
 \begin{equation}
     \begin{tikzcd}
         \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
                           & \dbase \arrow[u, "\dsection"', bend right]
     \end{tikzcd}
 \end{equation}
  such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$. The set of all global sections is denoted as $\Gamma(\dtotal)$. As illustrated in \autoref{fig:math:data:section}, the section is a continuous mapping from a location $\dbasepoint \in \dbase$ on the base space to a record $\delement \in \dfiber$ in the fiber. Assuming a trivial fiber bundle $\dtotal = \dbase \times \dfiber$, the section is 
 \begin{equation}
     \label{eq:section_return}
     \dsection(\dbasepoint) = (\dbasepoint, (g_{\dfiber_{0}}(\dbasepoint), \ldots, g_{\dfiber_{n}}(\dbasepoint)))
 \end{equation}
 where $g: \dbase \rightarrow \dfiber$ is the index function into the fiber component. This formulation of the section also holds on locally trivial sections of a non-trivial fiber bundle. As with \autoref{eq:math:data:fiber:decompose} and \autoref{eq:math:data:base:decompose}, \dsection\ can be decomposed into components 
 \begin{equation}
 \label{eq:math:data:section:decompose}
 \dsection= (\dsection_0,\ldots, \dsection_i, \dots, \dsection_n) 
 \end{equation}
 where each section $\dsection_i$ maps into a record on a component $\dfiber_i \in \dfiber$. This allows for accessing the data component wise in addition to accessing the data in terms of its location over \dbase.

@@

\subsubsection{Equivariant Topology: Graphic Base Space}
\label{sec:math:graphic:base}
\begin{figure}[htb]
  \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
  \includegraphics[width=\columnwidth]{retraction_maps.png}
  \caption{The 0D scatter \dbasepoint\ and 1D line \dbasepoint\ are thickened into \gbase\ with coordinates $\gbasepoint=(\gx, \gy)$ that are a region in an idealized 2D screen. The image has the same dimension in \gbase\ as in \dbase.}
  \label{fig:math:graphic:base}
 \end{figure}

Just as the \dbase\ encodes the connectivity of the records in the data, we propose an equivalent \gbase\ that encodes the connectivity of the rendered elements of the graphic. Formally, we require that \dbase\ be a deformation retract\cite{RetractionTopology2020} of \gbase\ so that \dbase\ and \gbase\ have the same homotopy. The surjective map $\vindex: \gbase \rightarrow \dbase$ 
\begin{equation}
    \begin{tikzcd}
        \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
        \dbase                   & \gbase \arrow[l, "\vindex"']
    \end{tikzcd}
\end{equation}
goes from region $\gbasepoint \in \gbase_{\dbasepoint}$ to its associated point $\gbasepoint$. While \gbase\ must have the same continuity as \dbase\, it is sometimes the  thickened version shown in \autoref{fig:math:graphic:base}. This thickening is necessary when the dimensionality of \dbase\ is less than the dimensionality of the target display. For example, a \dbasepoint\ that is 0D in \dbase\ cannot be represented on screen unless it is thickened to 2D to encode the connectivity of the points in \gfiber\ that visually represent the record at \dbasepoint.

@@

\subsubsection{Visual Component Maps}
\label{sec:math:artist:nu}
\begin{figure}[htb]
  \centering
  \includegraphics[width=\columnwidth]{equivariance_nu.png}
  \caption{In this artist, \vchannel\ maps the strings to the emojis. This \vchannel\ is equivariant because the monoid actions, represented by the colored arrows, are the same on the input \dsection\ and output \vsection\ sets.}
  \label{fig:math:artist:nu}
\end{figure}
 We define the visual transformers \vchannel\ 
\begin{equation}
  \label{eq:math:artist:nu}
  \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
as the set of equivariant maps $\vchannel_i: \dsection_i \mapsto \vsection_i$. Given $\monoid_i$ is the monoid action on $\dtotal_i$ and that there is a monoid ${\monoid_i}^{\prime}$ on $\vtotal_i$, then there is a monoid homomorphism from $\varphi:\monoid_i \rightarrow {\monoid_i}^{\prime}$ that \vchannel\ must preserve. As mentioned in \autoref{sec:math:data:monoid}, monoid actions define the structure on the fiber components and are therefore the basis for equivariance. 
A validly constructed \vchannel\ is one where the diagram of the monoid transform $m$ commutes such that 
\begin{equation}
  \label{eq:math:artist:nu_commute}
\begin{tikzcd}
  \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
  \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
\end{tikzcd}
\end{equation}
In general, the data fiber $\dfiber_{i}$ cannot be assumed to be of the same type as the visual fiber $\vfiber_{i}$ and the actions of \monoid\ on $\dfiber_{i}$ cannot be assumed to be the same as the actions of $\monoid^{\prime}$ on \vfiber; therefore an equivariant $\vchannel_i$ must satisfy the constraint  
\begin{equation}
\vchannel_i(m_{\delement}(\dtotal_i)) = \varphi(m_{\delement})(\vchannel_i(\dtotal_i))
\end{equation} 
such that $\varphi$ maps a monoid action on data to a monoid action on visual elements. However, we can construct a monoid action of \monoid\ on $\vfiber_{i}$ that is compatible with a monoid action of \monoid\ on $\dfiber_{i}$. We can compose the monoid actions on the visual fiber $\monoid^{\prime} \times \vfiber_{i} \rightarrow \vfiber_{i}$ with the homomorphism $\varphi$ that takes \monoid\ to $\monoid^{\prime}$. This allows us to define a monoid action on \vfiber\ of \monoid\ that is $(m, \velement) \rightarrow \varphi(m)\bullet\velement$. Therefore, without a loss of generality, we can assume that an action of \monoid\ acts on $\dfiber_{i}$ and on $\vfiber_{i}$ compatibly such that $\varphi$ is the identity function.

@@

\subsubsection{Visualization Assembly}
\begin{figure}[htb]
  \centering
  \includegraphics[width=\columnwidth]{path_of_q}
  \caption{$\vchannel_i$ functions convert data $\dsection_i$ to visual characteristics $\vsection_i$, then \vmark\ assembles $\vsection_i$ into a graphic $\gsection$ such that there is a map \vindex\ preserving the continuity of the data.} 
  \label{fig:math:artist:q}
\end{figure}p

As shown in \autoref{fig:math:artist:q}, \vchannel\ and \vmark\ are a map-reduce operation: map the data into their visual encodings \vsection\, reduce the encodings into a graphic \gsection. While it may seem intuitive that visualizations that generate the same graphic should do so consistently given the same input, we formalize this constraint such that it can be specified as a constraint of \vmark. We define the equivariant map as  $\vmark: \vsection \mapsto \gsection$ and an action on the subset of graphics $\vmark(\Gamma(\vtotal)) \in \Gamma(\gtotal)$ that \vmark\ can generate. We then define the constraint on \vmark such that if \vmark\ is applied to two visual sections $\vsection,\vsection^{\prime}$ that generate the same \gsection\, then the output of $\vsection,\vsection^{\prime}$ acted on by the same monoid $m$ must be the same.

\begin{figure}[htb]
  \centering
  \includegraphics[width=\columnwidth]{diff_type_q.png}
  \caption{These two glyphs are generated by the same \vmark\ function. The monoid 
  action $m_i$ on edge thickness $\vsection_i$ of the first glyph yields the thicker edge ${\vsection_i}^{\prime}$ in the second glyph.}
  \label{fig:math:arist:graphic}
\end{figure}

Lets call the visual representations of the components $\Gamma(\vtotal)=X$ and the graphic $\vmark(\Gamma(\vtotal))=Y$. If for elements of the monoid $m \in \monoid$ and for all $\vsection, \vsection^{\prime} \in X$, we define the monoid action on $X$ so that it is by definition equivariant
\begin{equation}
\vmark(\vsection) = \vmark(\vsection^{\prime})\implies \vmark(m\circ\vsection) = \vmark(m\circ\vsection^{\prime})
\end{equation}
then a monoid action on $Y$ can be defined as $m\circ \gsection = \gsection^{\prime}$. The transformed graphic $\gsection^{\prime}$ is equivariant to a transform on the visual bundle $\gsection^{\prime}=\vmark(m\circ \vsection)$ on a section that $\vsection \in \vmark^{-1}(\gsection)$ that must be part of generating \gsection. This equivariance is illustrated in figure \autoref{fig:math:artist:graphic}. Given fiber $\vfiber=(xpos,\, ypos,\, color,\, thickness)$, then sections $\vsection=(0,0,0,1)$ and $\vmark(\vsection) = \gsection$ generates a piece of the thin hollow circle. The action $m=(e, e, e, x+2)$, where e is identity, translates \vsection\ to  $\vsection^{\prime}=(e,e,e,3)$ and the corresponding action on \gsection\ causes $\vmark(\vsection^{\prime})$ to be the thicker circle in \autoref{fig:math:artist:graphic}.

We formally describe a glyph as \vmark\ applied to the regions \dbasepoint\ that map back to a set of path connected components $\dbasepath \subset \dbase$ as input:
\begin{equation}
\dbasepath = \{\dbasepathpoint \in \dbase \text{ s. t. } \exists \gamma \text{ s.t. } \gamma(0)=\dbasepoint \text{ and }\gamma(1)=\dbasepathpoint\}
\end{equation}
where the path\cite{ConnectedSpace2020}  $\gamma$ from \dbasepoint\ to \dbasepathpoint\ is a continuous function from the interval [0,1]. We define the glyph as the graphic generated by $\vmark(\gbase_{\dbasepathpoint})$
\begin{equation}
  \begin{tikzcd}
      \gtotal \arrow[r, shift left] & \gbase_\dbasepathpoint \arrow[rr, "\vindex(\gbasepoint)", shift left] \arrow[l, "\gsection(\gbase_\dbasepathpoint)"] &  & \dbasepath_{\dbasepoint} \arrow[ll, "\vindex^{-1}(\dbasepath)"]
      \end{tikzcd}
  \label{eq:mark}
\end{equation}
such that for every glyph there is at least one corresponding region on \dbase, in keeping with the definition of glyph as any differentiable element put forth by Ziemkiewicz and Kosara\cite{ziemkiewiczEmbeddingInformationVisualization2009}. The primitive point, line, and area marks\cite{bertinSemiologyGraphicsDiagrams2011a,carpendaleVisualRepresentationSemiology} are specially cased glyphs.

\begin{figure}
  \centering
  \includegraphics{}
\end{figure}

In \autoref{fig:teaser}, we illustrate the output of a minimal Q that will generate distinguishable graphical marks: non-overlapping scatter points, a non-infinitely thin line, and an image.

The scatter plot can be defined as $\vmark(xpos, ypos)(\alpha, \beta)$ where color $\gsection_{RGB} = (0,0,0)$ is defined as part of \vmark\ and $\gbasepoint=(\alpha, \beta)$ defines the region on \gbase.

The position of this swatch of color can be computed relative to the location on the disc $\gbase_{\dbasepoint}$ as shown in figure~\ref{fig:artist_scatter}:
\begin{align}
x &= size *\alpha \cos(\beta) + xpos \\
y &= size *\alpha \sin(\beta) + ypos
\end{align}

such that $\gsection(\gbasepoint) = (x, y, 0, 0, 0)$ colors the point (x,y) black. Here $size$ can either be defined inside \vmark\ or it could also be a parameter in \vtotal\ that is passed along with (xpos, vpos). As seen in figure~\ref{fig:artist_scatter}, a scatter has a direct mapping from a region on $\gbase_{k}$ to its corresponding \dbasepoint.
